---
title: "ATPA 2.3 - Data Transformation and Cleaning"

---

CHUNK 1: Load the tidyverse and create the data frame
```{r}
library(tidyverse)
trials <- data.frame(
  Phase = rep(c("A", "B", "C"), each = 3),
  Count = c(50, 40, 65, 70, 70, 90, 30, 20, 25),
  Successes = c(20, 17, 31, 40, 33, 41, 12, 9, 9),
  Location = c("Austin Texas", "Nashville Tennessee", "Nashville Tennessee", "Dallas Texas", "Seattle Washington", "Houston Texas", "Orlando Florida", "Cincinnati Ohio", "El Paso Texas")
)
trials$Phase <- as.factor(trials$Phase)
trials
```

CHUNK 2: Subset to select Phase = A
```{r}
# Subset to only Phase A
trials_A <- trials %>% filter(Phase == "A")
trials_A
```

CHUNK 2A: Subsetting using brackets
```{r}
trials_A <- trials[trials$Phase == "A", ]
trials_A
```

CHUNK 3: Subset to select Phase = A or B
```{r}
# Either Phase A or B
trials_ABv1 <- trials %>% filter(Phase %in% c("A", "B")) # or
trials_ABv2 <- trials %>% filter(Phase == "A" | Phase == "B")
trials_ABv1
trials_ABv2
```

CHUNK 4: Subset to select Phase not equal to A
```{r}
# Not Phase A
trials_NotAv1 <- trials %>% filter(Phase != "A")
trials_NotAv2 <- trials %>% filter(!(Phase == "A"))
trials_NotAv1
trials_NotAv2
```

CHUNK 5: Exercise 2.3.1: Use three different ways to subset to all locations where Phase is A or C
```{r}

```

CHUNK 6: Exercise 2.3.1: Solution with four ways presented
```{r}
sol1 <- trials %>% filter(Phase == "A" | Phase == "C")
sol2 <- trials %>% filter(Phase != "B")
sol3 <- trials %>% filter(!(Phase == "B"))
sol4 <- trials %>% filter(Phase %in% c("A", "C"))
sol1
sol2
sol3
sol4
```

CHUNK 7: Subset to Counts greater than or equal to 50
```{r}
# Count greater than or equal to 50
trials_50 <- trials %>% filter(Count >= 50)
trials_50
```

CHUNK 8: Subset to Count less than 40 or greater than 70
```{r}
# Count less than 40 or greater than 70
trials_4070 <- trials %>% filter(Count < 40 | Count > 70)
trials_4070
```

CHUNK 9: Subset to Count between 45 and 80
```{r}
# Count between 45 and 80
trials_4580 <- trials %>% filter(Count > 45 & Count < 80)
trials_4580
```

CHUNK 10: Subset to Location contains Texas
```{r}
# Location contains Texas
trials_Texas <- trials %>% filter(str_detect(Location, "Texas"))
trials_Texas
```

CHUNK 11: Exercise 2.3.2: Subset to locations as indicated
```{r}
# Subset to locations that
# a) contain "n"


# b) ends with "n"


# c) start with "n"
```

CHUNK 12: Exercise 2.3.2: Solution
```{r}
# Subset to locations that
# a) contain "n"
trials_contain_n <- trials %>% filter(str_detect(Location, "n"))
trials_contain_n

# b) ends with "n"
trials_end_n <- trials %>% filter(str_detect(Location, "n$"))
trials_end_n

# c) start with "n"
trials_start_n <- trials %>% filter(str_detect(Location, "^n"))
trials_start_n
```

CHUNK 13: Subset to Phase is A and Count exceeds 40
```{r}
# Phase is A and count is more than 40
trials_A40 <- trials %>% filter(Phase == "A" & Count > 40)
trials_A40
```

CHUNK 14: Exercise 2.3.3: Subset to the conditions as outlined 
```{r}

# Subset to observations where Count is less than 50 AND Successes are more than 15.


# Subset to observations where either Count is less than 50 OR Successes are more than 15.
```

CHUNK 15: Exercise 2.3.3: Solution
```{r}

# Subset to observations where Count is less than 50 AND Successes are more than 15.
trials_50and15 <- trials %>% filter(Count < 50 & Successes > 15)
trials_50and15

# Subset to observations where either Count is less than 50 OR Successes are more than 15.
trials_50or15 <- trials %>% filter(Count < 50 | Successes > 15)
trials_50or15
```

CHUNK 16: Exercise 2.3.4: Subset to include only locations that include an “s” AND have more than 10 successes
```{r}

```

CHUNK 17: Exercise 2.3.4: Solution
```{r}

trials_sand10 <- trials %>% filter(str_detect(Location, "s") & Successes > 10)
trials_sand10
```

CHUNK 18: Subset to the Count variable
```{r}
# Subset to the Count variable
trials_Count <- trials %>% select(Count)
trials_Count
```

CHUNK 19: Subset to the Count and Successes variables
```{r}
# Subset to counts and successes
trials_CandSv1 <- trials %>% select(Count, Successes) # or
trials_CandSv2 <- trials %>% select(2:3)
trials_CandSv1
trials_CandSv2
```

CHUNK 20: Subset to delete the variable Location
```{r}
# Subset to all variables except Location
trials_noLocation <- trials %>% select(-Location)
trials_noLocation
```

CHUNK 21: Exercise 2.3.5: Subset to include Count and Successes for those in Phase A
```{r}

```

CHUNK 22: Exercise 2.3.5: Solution
```{r}
sol1 <- trials %>%
  filter(Phase == "A") %>%
  select(2:3) # Note that if select is done before filter an error results. That is because after the selection there is no longer a Phase variable.
sol2 <- trials[trials$Phase == "A", c("Count", "Successes")]
sol1
sol2
```

CHUNK 23: Order the data by increasing count
```{r}
# Sort by Count, low to high
trials_sortCountInc <- trials %>% arrange(Count)
trials_sortCountInc
```

CHUNK 24: Order the data by decreasing count
```{r}
# Sort counts high to low
trials_sortCountDec <- trials %>% arrange(desc(Count))
trials_sortCountDec
```

CHUNK 25: Order the data by increasing count, using increasing successes to break ties
```{r}
# Successes is a tie breaker for counts
trials_tiebreak1 <- trials %>% arrange(Count, Successes)
trials_tiebreak1

# The following breaks the tie using decreasing successes
trials_tiebreak2 <- trials %>% arrange(Count, desc(Successes))
trials_tiebreak2
```

CHUNK 26: Create the new variable Rate = Successes/Count
```{r}
# Create a new variable
trials_rate1 <- trials %>% mutate(Rate = Successes / Count)
trials_rate1

# This can also be done directly
trials_rate2 <- trials
trials_rate2$Rate <- trials_rate2$Successes / trials_rate2$Count
trials_rate2
```

CHUNK 27: Creating two variables at once
```{r}
# Create multiple variables
trials_multi <- trials %>% mutate(Large_Counts = Count > 50, In_Texas = str_detect(Location, "Texas"))
trials_multi
```

CHUNK 28: Create bins for the Count variable
```{r}
# Create binned variable
trials_bin <- trials %>% mutate(Count_Level = cut(Count, breaks = c(0, 35, 65, 100), labels = c("Low", "Medium", "High")))
trials_bin
```

CHUNK 29: Overwrite Location by making each observation have the value "Unknown"
```{r}
# Reassign an existing variable
trials_unknown <- trials %>% mutate(Location = "Unknown")
trials_unknown
```

CHUNK 30: Exercise 2.3.6: Create two sets of bins for the Successes variable
```{r}
# (a) Bin Successes using cutoffs of 15 and 30


# (b) Bin Successes into four bins with an approximately equal number of observations in each bin. Check help on cut.
```

CHUNK 31: Exercise 2.3.6: Solution
```{r}
# (a) Bin Successes using cutoffs of 15 and 30
trials_bin1 <- trials %>% mutate(Success_Level1 = cut(Successes, breaks = c(0, 15, 30, 100), labels = c("Low", "Medium", "High")))
trials_bin1

# (b) Bin Successes into four bins with an approximately equal number of observations in each bin.
trials_bin2 <- trials %>% mutate(Success_Level2 = cut(Successes, breaks = 4)) # By not specifying the labels, they will be generated automatically. See help to learn why the lower limit of the first interval was changed.
trials_bin2
```

CHUNK 32: Obtain the mean count by phase
```{r}
# Find mean counts by phase
trials_means <- trials %>%
  group_by(Phase) %>%
  summarize(AverageByPhase = mean(Count))
trials_means
```

CHUNK 33: Obtain the number of observations in each phase
```{r}
# Counts how many in each phase
trials_number <- trials %>%
  group_by(Phase) %>%
  summarize(NumPhase = n())
trials_number
```

CHUNK 34: Obtain the number of unique locations in each phase
```{r}
# Count the number of unique locations in each phase
trials_unique <- trials %>%
  group_by(Phase) %>%
  summarize(NumLocs = n_distinct(Location))
trials_unique
```

CHUNK 35: Exercise 2.3.7: Obtain the number of unique values of Count in each Phase
```{r}
# Count the number of unique Count values there are in each Phase.
```

CHUNK 36: Exercise 2.3.7: Solution
```{r}
# Count the number of unique Count values there are in each Phase.
trials_countOfCount <- trials %>%
  group_by(Phase) %>%
  summarize(NumCounts = n_distinct(Count))
trials_countOfCount
```

CHUNK 37: Exercise 2.3.8: Before doing this, you may want to clear the environment as we are now finished with the trials data. You can use the broom in the Environment area or run the command rm(list = ls()) in the Console
```{r}
# install.packages(nycflights13) # Uncomment if you need to install this package
library(nycflights13)
# While it doesn't appear in the environment, an object called flights is available that has the data. For example, summary(flights) allows you to see a summary of the data.

# 1. Filter observations to only values where carrier is equal to “UA”

# 2. Filter observations to only where the dest variable is either “MIA” or “ORD”

# 3. Filter observations to only where air_time is between 60 and 120 minutes.

# 4. Filter variables to only sched_dep_time and dep_delay

# 5. Order the data from longest distance to shortest distance

# 6. Create a new variable called delayed that returns TRUE if dep_delay is greater than 0 and FALSE otherwise.

# 7. Create a new variable called mph that is equal to distance divided by air_time times 60

# 8. Determine mean dep_delay values by carrier. Call it mean_delay

# 9. Each plane has a unique value for tailnum. Determine how many distinct planes each carrier uses. Call it num_planes.
```

CHUNK 38: Exercise 2.3.8: Solution
```{r}
# install.packages(nycflights13)
library(nycflights13)


# 1. Filter observations to only values where carrier is equal to “UA”
flights %>%
  filter(carrier == "UA") # In this solution the results are not put in a new object, rather they are just displayed. Because the flights is a tibble and not a dataframe only the first 10 rows will be displayed.

# 2. Filter observations to only where the dest variable is either “MIA” or “ORD”
flights %>%
  filter(dest %in% c("MIA", "ORD"))

# 3. Filter observations to only where air_time is between 60 and 120 minutes.
flights %>%
  filter(air_time > 60 & air_time < 120)

# 4. Filter variables to only sched_dep_time and dep_delay
flights %>%
  select(sched_dep_time, dep_delay)

# 5. Order the data from longest distance to shortest distance
flights %>%
  arrange(desc(distance))

# 6. Create a new variable called delayed that returns TRUE if dep_delay is greater than 0 and FALSE otherwise.
flights %>%
  mutate(delayed = dep_delay > 0)

# 7. Create a new variable called mph that is equal to distance divided by air_time times 60
flights %>%
  mutate(mph = distance / air_time * 60)

# 8. Determine mean dep_delay values by carrier. Call it mean_delay
flights %>%
  group_by(carrier) %>%
  summarize(mean_delay = mean(dep_delay, na.rm = TRUE))

# 9. Each plane has a unique value for tailnum. Determine how many distinct planes each carrier uses. Call it num_planes.
flights %>%
  group_by(carrier) %>%
  summarize(num_planes = n_distinct(tailnum))
```

CHUNK 39: Exercise 2.3.9
```{r}
# Determine the top 5 most common destinations from JFK airport.
```

CHUNK 40: Exercise 2.3.9: Solution
```{r}
# Determine the top 5 most common destinations from JFK airport.
flights %>%
  filter(origin == "JFK") %>%
  group_by(dest) %>%
  summarize(dest_count = n()) %>%
  arrange(desc(dest_count)) %>%
  filter(row_number() < 6)
```

CHUNK 41: Exercise 2.3.10
```{r}
# Create a new variable that is the total number of nights booked by adding stay_in_weekday_nights and stay_in_weekend_nights. Keep only bookings where the reserved_room_type and the assigned_room_type are equal. Find the mean number of total nights stayed grouped by assigned_room_type.
```

CHUNK 42: Exercise 2.3.10: Solution
```{r}
# Create a new variable that is the total number of nights booked by adding stay_in_weekday_nights and stay_in_weekend_nights. Keep only bookings where the reserved_room_type and the assigned_room_type are equal. Find the mean number of total nights stayed grouped by assigned_room_type.

hotel_bookings <- read_csv("hotel_bookings.csv")
hotel_bookings %>%
  mutate(total_stay = stays_in_week_nights + stays_in_weekend_nights) %>%
  filter(assigned_room_type == reserved_room_type) %>%
  group_by(assigned_room_type) %>%
  summarize(mean_stay = mean(total_stay))
```

CHUNK 43: Create and display the sample data
```{r}
library(tidyverse)
security <- tibble(
  "Camera" = c("Lobby", "Main Floor", "Back Door", "Front Door", "Offices"),
  "Total Hours Recorded" = c(51, 59, -1, 22, 27),
  "Incidents" = c(2, 1, 0, 0, NA)
)
security
```

CHUNK 44: Rename Total Hours Recorded to Hours_Recorded
```{r}
# Renaming variables
security <- security %>% rename(Hours_Recorded = "Total Hours Recorded") # Putting quotation marks around the variable name is not normally required, but is necessary in this case due to the spaces in the old name.
security
```

CHUNK 45: Remove observations with Hours_Recorded = -1 or Incidents = NA
```{r}
# Remove missing values specifically for each row
security %>% filter(Hours_Recorded != -1, !is.na(Incidents)) # This time we are not saving the output as we may decide on a different way to deal with missing values.
```

CHUNK 46: Remove observations with NAs without filtering
```{r}
# Remove all instances of NA
security %>% na.omit()
```

CHUNK 47: Change all instances of -1 to NA
```{r}
# CHUNK 5
# Recode all instances of missing to NA
# Because this code makes a permanent change to the data, we do it with a new version
security_recode <- security
security_recode[security_recode == -1] <- NA
security_recode
rm(security_recode)
```

CHUNK 48: Exercise 2.3.11: Remove observations with 0 -1, and NA as if they are all missing values. Rename Incidents to Occurrences
```{r}

```

CHUNK 49: Exercise 2.3.11: Solution
```{r}
# Remove observations with 0 -1, and NA as if they are all missing values. Rename Incidents to Occurrences
security %>%
  filter(
    Hours_Recorded != -1,
    !is.na(Incidents),
    Incidents != 0
  ) %>%
  rename(Occurrences = Incidents)
```

CHUNK 50: String to Numeric
```{r}
# String to Numeric
x <- "3"; y <- "4"
as.numeric(x) + as.numeric(y)
```

CHUNK 51: Numeric to String
```{r}
# Numeric to String
as.character(3)
```

CHUNK 52: Changing case
```{r}
# Case
str_to_lower("String")
str_to_upper("String")
```

CHUNK 53: Working with a vector of strings
```{r}
X <- c("String1", "String2")
str_to_lower(X)
str_to_upper(X)
```

CHUNK 54: Replacing string content
```{r}
# Replacement
x <- "my_string"
str_replace_all(x, "my", "your")
X <- c("my_string1", "my_string2")
str_replace_all(X, "my", "your")
```

CHUNK 55: Replacing string content
```{r}
# Converting money to numeric
x <- "$84,510"
as.numeric(str_replace_all(x, "[$,]", ""))
```

CHUNK 56: Obtaining substrings
```{r}
# Substrings
x <- "apple"
str_sub(x, 1, 3)
X <- c("apple", "pear")
str_sub(X, 1, 3)
```

CHUNK 57: Combining and splitting strings
```{r}
# Combining Strings
a <- "apple"
c <- "crisp"
str_c(a, " ", c) # apple crisp

str_c(a, c) # There is no space between the words here
h <- "honey"
str_c(c(h, a, c), collapse = "") # no space between the words
str_c(c(h, a, c), collapse = "-") # places a dash between the words
str_c(c(h, a, c), collapse = " ") # places a space between the words

# Splitting Strings
dessert <- "honey apple crisp"
str_split(dessert, " ", simplify = TRUE)
```

CHUNK 58: Exercise 2.3.12: Perform the indicated tasks on a vector of words
```{r}

my_words <- c("Actuarial", "Exam", "Passed")
# 1. Change all the letters to lower case


# 2. Replace each letter “a” with a dollar sign, “$”


# 2. Combine all three words putting a space in between

```

CHUNK 59: Exercise 2.3.12: Solution
```{r}

my_words <- c("Actuarial", "Exam", "Passed")
# 1. Change all the letter to lower case
str_to_lower(my_words)

# 2. Replace all “a”’s with dollar signs, “$”
str_replace_all(my_words, "a", "$")

# 3. Combine all three words putting a space in between
str_c(my_words, collapse = " ")
```

CHUNK 60: Making date variables
```{r}
# install.packages(lubridate) # Uncomment if this package needs to be installed
library(lubridate)
make_date(year = 2022, month = 1, day = 1)
make_datetime(year = 2022, month = 1, day = 1, hour = 9, min = 30, sec = 55)
```

CHUNK 61: Converting string to date
```{r}
Date1 <- "10/01/2021"
Date2 <- "20211001"
mdy(Date1)
ymd(Date2)
Date_Time <- "10-01-2021 5:30:41"
mdy_hms(Date_Time)
```

CHUNK 62: Calculating time differences
```{r}
# Time differences
today() - make_date(2021, 01, 01)
now() - make_datetime(2021, 01, 01, 9, 30, 55)
Time1 <- hms("12:30:09")
Time2 <- hms("12:35:30")
Time2 - Time1
```

CHUNK 63: Extract day/month/year from a date variable
```{r}
my_date <- mdy("07102003")
month(my_date) # Extracts the month as a number
month(my_date, label = TRUE) # Extracts the name of the month
year(my_date) # Extracts the year
mday(my_date) # Extracts the day of the month
wday(my_date) # Extracts the day of the week as a number, Sunday is 1
wday(my_date, label = TRUE) # Extracts the day of week as a name
```

CHUNK 64: Exercise 2.3.13: Convert the Year, Month, and Day variables in the flight data to a single date variable
```{r}
library(nycflights13)

```

CHUNK 65: Exercise 2.3.13: Solution
```{r}
library(nycflights13)
flights <- flights %>%
  mutate(Date = make_date(year, month, day))
flights[1:10, "Date"]
```

CHUNK 66: Create ordered and unordered factor variables
```{r}
# Create an ordered factor variable
fct_ordered <- factor(c("good", "good", "poor", "okay"), levels = c("poor", "okay", "good"), ordered = TRUE)
fct_ordered

# Create an unordered factor variable
fct_unordered <- factor(c("good", "good", "poor", "okay"), ordered = FALSE)
fct_unordered

# Convert an ordered factor variable to unordered
fct_unordered_2 <- factor(fct_ordered, ordered = FALSE)
fct_unordered_2
```

CHUNK 67: Factor recoding and combining
```{r}
# Survey data set, note that factor() must be used else these will be interpreted as character variables. The display confirms that Sex and Answer are factor variables. The display shows the variety of levels observed.
survey <- tibble(
  "Individual" = 1:5,
  "Sex" = factor(c("Male", "male", "fem.", "Female", "M")),
  "Answer" = factor(c("Y", "Y", "N", "Yes", "n"))
)
survey
levels(survey$Sex)
levels(survey$Answer)
```

CHUNK 68: Change factors to upper case
```{r}
# Recoding
survey_upper <- survey %>% mutate(Sex = factor(str_to_upper(Sex)), Answer = factor(str_to_upper(Answer)))
survey_upper
levels(survey_upper$Sex)
levels(survey_upper$Answer)
```

CHUNK 69: Change factors to upper case and retain only first letter
```{r}
survey_one <- survey %>% mutate(Sex = factor(str_sub(str_to_upper(Sex), 1, 1)), Answer = factor(str_sub(str_to_upper(Answer), 1, 1)))
survey_one
levels(survey_one$Sex)
levels(survey_one$Answer)
```

CHUNK 70: Change factor values with direct recoding
```{r}
# Recoding directly
survey_direct <- survey %>% mutate(Sex = fct_recode(Sex, Male = "M", Male = "male", Female = "fem."))
survey_direct
levels(survey_direct$Sex)
```

CHUNK 71: Combining factors
```{r}
# Factor combining
library(nycflights13)
small_levels <- c("AS", "F9", "HA", "OO", "YV")
flights_combined <- flights %>%
  mutate(carrier = factor(ifelse(carrier %in% small_levels, "Other", carrier))) # Carrier is imported as a character variable. Here we convert it to a factor variable.
levels(flights_combined$carrier)
table(flights_combined$carrier)
```

CHUNK 72: Converting Boolean to integer
```{r}
# Convert Boolean to integer
security <- tibble(
  "Camera" = c("Lobby", "Main Floor", "Back Door", "Front Door", "Offices"),
  "Total Hours Recorded" = c(51, 59, -1, 22, 27),
  "Incidents" = c(2, 1, 0, 0, NA)
)
security[, 2] > 30 # To create a Boolean variable for this example
(security[, 2] > 30) * 1 # To convert it to 0 and 1
```

CHUNK 73: Making a factor variable numeric
```{r}
# Try converting factors to numeric
my_factor <- as.factor(c(15, 5, 10, 15, 15, 15, 10))
my_factor
as.numeric(my_factor) # Internally, R codes factor variables as 1, 2, etc. with a table that assigns factor values. In this case 5 is assigned level 1, 10 is assigned level 2, and 15 is assigned level 3. The conversion to numeric uses 1, 2, and 3 and not 5, 10, and 15.

# First convert the factor to character, then to numeric
my_char <- as.character(my_factor)
my_num <- as.numeric(my_char)
my_num
```

CHUNK 74: Exercise 2.3.14
```{r}
Auto_Names <- c("symboling", "normalized_losses", "make", "fuel_type", "aspiration", "num_doors", "body_style", "drive_wheels", "engine_location", "wheel_base", "length", "width", "height", "curb_weight", "engine_type", "num_cylinders", "engine_size", "fuel_system", "bore", "stroke", "compression_ratio", "horsepower", "peak_rpm", "city_mpg", "highway_mpg", "price")

auto <- read_csv("automobile.csv", col_names = Auto_Names)

# 1 Replace those “?” values with NAs


# 2 Remove all records with missing values


# 3 Rename the variable drive_wheels to drive_type


# 4 Recode all the variables in the new variable drive_type. Make “rwd” into “rear”, “fwd” into “front”, and “4wd” into “four”


# 5 Change the values in the variable engine_location so that only the first two letters show and make those two letters all uppercase


# 6 There are some numeric columns that had “?”, and as a result, they were marked as strings or characters. Convert those variables to numeric


# 7 Combine factors in the body_style variable so that there are only three factors: “sedan”, “hatchback”, and “Other”
```

CHUNK 75: Exercise 2.3.14: Solution
```{r}
Auto_Names <- c("symboling", "normalized_losses", "make", "fuel_type", "aspiration", "num_doors", "body_style", "drive_wheels", "engine_location", "wheel_base", "length", "width", "height", "curb_weight", "engine_type", "num_cylinders", "engine_size", "fuel_system", "bore", "stroke", "compression_ratio", "horsepower", "peak_rpm", "city_mpg", "highway_mpg", "price")

auto <- read_csv("automobile.csv", col_names = Auto_Names)

# 1 Replace those “?” values with NAs
auto[auto == "?"] <- NA

# 2 Remove all records with missing values
auto <- na.omit(auto)

# 3 Rename the variable drive_wheels to drive_type.
auto <- auto %>% rename("drive_type" = "drive_wheels")

# 4 Recode all the variables in the new variable drive_type. Make “rwd” into “rear”, “fwd” into “front”, and “4wd” into “four”.
auto <- auto %>% mutate(drive_type = fct_recode(drive_type, rear = "rwd", front = "fwd", four = "4wd"))

# 5 Change the values in the variable engine_location  so that only the first two letters show and make those two letters all uppercase.
auto <- auto %>%
  mutate(engine_location = str_to_upper(str_sub(engine_location, 1, 2)))

# 6 There are some numeric columns that had “?”, and as a result, they were marked as strings or characters. Convert those variables to numeric.
auto <- auto %>%
  mutate(
    normalized_losses = as.numeric(normalized_losses),
    stroke = as.numeric(stroke),
    bore = as.numeric(bore),
    price = as.numeric(price)
  )

# 7 Combine factors in the body_style variable so that there are only three factors: “sedan”, “hatchback”, and “Other”
auto <- auto %>%
  mutate(body_style = ifelse(body_style %in% c("sedan", "hatchback"), body_style, "Other"))

auto
```
