---
title: "Module 2 Section 7 Case Study 2"
---

Read in questionnaire data
```{r}
library(tidyverse)
library(XML)
chiro <- read_csv("questionnaire.csv")
```

Read in visits data, extract variables, and create data frame
```{r}
# read xml structure
visits <- xmlParse("visits.xml")

# extract the variables
ID <- xpathSApply(visits, path = "/REPORT/records/visit/ID", xmlValue)
type <- xpathSApply(visits, path = "/REPORT/records/visit/type", xmlValue)
reschedule <- xpathSApply(visits, path = "/REPORT/records/visit/reschedule", xmlValue)
year <- xpathSApply(visits, path = "/REPORT/records/visit", xmlGetAttr, name = "year")

# combine the variables into a data frame
all_visits <- tibble(ID, type, reschedule, year)
```

Validate the year variable
```{r}
visits_year <- all_visits %>%
  group_by(ID) %>%
  summarize("first_visit" = min(year))
quest_year <- chiro %>%
  select(ID, year) %>%
  mutate(ID = as.character(ID))
validate <- inner_join(visits_year, quest_year)
sum(validate$first_visit != validate$year)
```

Aggregate the visits data, count the number of visits in 2018
```{r}
agg_2018 <- all_visits %>%
  filter(year == 2018) %>%
  group_by(ID) %>%
  summarize(counts_2018 = n())

agg_2018 %>% ggplot(aes(counts_2018)) +
  geom_histogram() +
  labs(
    title = "Number of Chiropractic Visits in 2018",
    x = "Visits", y = "Count"
  )
```

Aggregate the visits data for 2017 for predictor variables
```{r}
agg_2017 <- all_visits %>%
  filter(year == 2017) %>%
  group_by(ID) %>%
  summarize(
    counts_2017 = n(),
    perc_ext = mean(type == "extended"),
    perc_resc = mean(reschedule == "yes")
  )

agg_all <- inner_join(agg_2018, agg_2017, by = "ID")
```

Create the variables for the causes
```{r}
chiro2 <- chiro %>%
  mutate(
    pain = str_detect(why_visit, "pain"),
    stiffness = str_detect(why_visit, "stiffness"),
    recovery = str_detect(why_visit, "recovery"),
    functionality = str_detect(why_visit, "functionality")
  ) %>%
  select(-why_visit)
```

Fix missing values in past_surgery
```{r}
table(chiro2$year, chiro2$past_surgery, useNA = "ifany")

# Remove the variable
chiro2$past_surgery <- NULL
```

Lump small factor levels of how_know together
```{r}
sort(table(chiro2$how_know))
chiro3 <- chiro2 %>% mutate(how_know = fct_lump_min(how_know, 20))
```

Combine the visits and questionnaire data together
Change non-ID character variable to a factor variable
```{r}
chiro3$ID <- as.character(chiro3$ID) # this is very important
# Because XML data is read in as all characters but csv
# will be readin as numeric, have to change one or the other
chiro_final <- inner_join(chiro3, agg_all, by = "ID")
chiro_final$sex <- as.factor(chiro_final$sex)
chiro_final
summary(chiro_final[, 1:4])
summary(chiro_final[, 5:8])
summary(chiro_final[, 9:12])
summary(chiro_final[, 13:14])
```
