---
title: "ATPA 2.2 - Reading and Writing Data"
---

CHUNK 1: Creates a tibble (a special data frame) from existing variables
```{r}
library(tidyverse)
Var1 <- c(1,2,3)
Var2 <- c("a","b","c")
Df <- tibble("Numbers" = Var1, "Letters" = Var2)
Df
```

CHUNK 2: Adds an additional variable to a data frame
```{r}
Var3 <- c(TRUE, FALSE, FALSE)
Df$Boolean = Var3 # adds Var3 and names it Boolean
Df
```

CHUNK 3: Creates a list
```{r}
# A list called food
food <- list("breakfast" = c("eggs","yogurt","cereal"),
             "lunch" = c("sandwich","soup"),
             "dinner" = c("pasta","steak","pizza","fish"))

# Extract the first element of the list
food[[1]]
food[["breakfast"]]

```


CHUNK 4: Reads the football_space.txt file into a data frame
```{r}
# Space delimited file
football1 <- read_table("football_space.txt")
football1
```

CHUNK 5: Reads in football_comma.csv file into a data frame
```{r}
# comma delimited file (csv)
football2 <- read_csv("football_comma.csv")
football2
```

CHUNK 6: Reads in football_semicolon.txt file into a data frame
```{r}
# comma delimited file (csv)
football3 <- read_delim("football_semicolon.txt",delim=";")
football3
```

CHUNK 7: Reads in football_skip.csv file into a data frame while skipping the first 4 lines
```{r}
# File with non-data as introduction
football4 <- read_csv("football_skip.csv",skip = 4)
football4
```

CHUNK 8: Reads in football_nohead.csv and adds column names
```{r}
# File with no header, provide column names
football5 <- read_csv("football_nohead.csv",col_names = c("Game","Win","Keeper"))
football5

# Read in with no header and add variable names after
football5 <- read_csv("Football_nohead.csv",col_names = FALSE)
names(football5) = c("Game","Win","QB")
```

CHUNK 9: Exercise 2.2.1: Read in automobile.csv and assign the labels given
```{r}
Auto_Names <- c("symboling", "normalized_losses", "make", "fuel_type,aspiration", "num_doors", "body_style", "drive_wheels","engine_location", "wheel_base", "length", "width", "height", "curb_weight", "engine_type","num_cylinders", "engine_size", "fuel_system", "bore", "stroke", "compression_ratio", "horsepower", "peak_rpm", "city_mpg", "highway_mpg", "price")


```

CHUNK 10: Exercise 2.2.1: Solution
```{r}
Auto_Names <- c("symboling", "normalized_losses", "make", "fuel_type","aspiration", "num_doors", "body_style", "drive_wheels","engine_location", "wheel_base", "length", "width", "height", "curb_weight", "engine_type","num_cylinders", "engine_size", "fuel_system", "bore", "stroke", "compression_ratio", "horsepower", "peak_rpm", "city_mpg", "highway_mpg", "price")

automobile <- read_csv("automobile.csv",col_names = Auto_Names)

# View the data set
automobile %>% View
```

CHUNK 11: Reading simple XML file into a data frame
```{r}
#install.packages(“XML”) # Run this only once to get the library installed
library(XML)
# View the file contents in XML format
test_xml <- xmlParse("test_simple.xml")
test_xml

# Read the data and convert to a data frame
test_df <- xmlToDataFrame("test_simple.xml")
test_df

```

CHUNK 12:
```{r}
# Read in xml file
test <-  xmlParse("test_one_student.xml")
# Locate the scores element
scores <- getNodeSet(test,path="/Test_Results/Candidate/Scores")
# Convert to data frame. Note that double brackets [[1]] must be used on scores because the output was a list
xmlToDataFrame(scores[[1]])
```

CHUNK 13: Read in xml data with multiple entries
```{r}
# Read in xml file
test <-  xmlParse("test_two_students.xml")
# Locate the scores element
scores <- getNodeSet(test,path="/Test_Results/Candidate/Scores")
# scores is now a list where each element is a different candidate
Steve_scores <- xmlToDataFrame(scores[[1]])
Steve_scores
Chalise_scores <- xmlToDataFrame(scores[[2]])
Chalise_scores
```

CHUNK 14: Extracting attributes and element values to create a data frame
```{r}
# Read xml file
test_xml <- xmlParse("test_one_student.xml")
# Extract the Points values
points <- xpathSApply(test_xml, path = "/Test_Results/Candidate/Scores/Question/Points", xmlValue)

# Extract the bank attribute
bank <- xpathSApply(test_xml, path = "/Test_Results/Candidate/Scores/Question", xmlGetAttr,name="bank")

# Combine into data frame
xml_df <- tibble("bank" = bank,"points" = points)
xml_df

```

CHUNK 15: Extracting attributes and element values from a mortality xml file
```{r}
mort <- xmlParse("mort.xml")
# Extract age from the attribute t
age <- xpathSApply(mort,path="/XTbML/Table/Values/Axis/Y",xmlGetAttr,name="t")

# Extract mortality rate from the Y elements
mortality <- xpathSApply(mort,path="/XTbML/Table/Values/Axis/Y",xmlValue)

# Combine into data frame
mort_df <- tibble("age" = age, "mortality" = mortality)
mort_df
```

CHUNK 16: Exercise 2.2.2: Extract the attributes from the author elements
```{r}

```

CHUNK 17: Exercise 2.2.2: Solution
```{r}
dis <- xmlParse("disablement.xml")
# Extract age from the attribute t
age <- xpathSApply(dis,path="/XTbML/Table/Values/Axis/Y",xmlGetAttr,name="t")

# Extract mortality rate from the Y elements
disablement <- xpathSApply(dis,path="/XTbML/Table/Values/Axis/Y",xmlValue)

# Combine into data frame
dis_df <- tibble("age" = age, "disablement" = disablement)
dis_df
```

CHUNK 18: Reading simple JSON file into a data frame
```{r}
# install.packages("jsonlite") # Run this only once
library(jsonlite)

# Read in the json data file
test_df1 <- fromJSON("test_simple.json")
test_df1
```

CHUNK 19: Read nested JSON file and unnest
```{r}
# Read in the nested data file and unnest the nested structure
fromJSON("test_nested.json") %>% unpack(results)

```

CHUNK 20: Read nested JSON file and unnest the structure
```{r}
# unnest the structure using the unnest function because the Topics variable has square brackets
fromJSON("test2_nested.json") %>% unnest(results)
```

CHUNK 21: Read nested JSON file and unpack the structure
```{r}
# Read in json file and get test variable, then unnest results variable
fromJSON("test3_nested.json")[["test"]] %>% unnest(results)
```

CHUNK 22: Exercise 2.2.3 Read in and create a data frame from colors.json file
```{r}

```

CHUNK 23: Exercise 2.2.3: Solution
```{r}
color_nested <- fromJSON("colors.json")
color_nested[["colors"]] %>%  unpack(code)
```

CHUNK 24: Exercise 2.2.4: Read in the menu.xls file into three dataframes, Dinner, Sides, and Desserts
```{r}
# Install the readxl package
# install.packages("readxl")
library(readxl)


```

CHUNK 25: Exercise 2.2.4: Solution
```{r}
# Install the readxl package
# install.packages("readxl")
library(readxl)
Dinner <- read_excel("menu.xls",sheet=1)
Sides <- read_excel("menu.xls",sheet=2)
Desserts <- read_excel("menu.xls",sheet=3)
head(Dinner)
head(Sides)
head(Desserts)
```

CHUNK 26: Create a tall data frame
```{r}
#Creates tall data frame
tall_books <- data.frame(
  "ID" = c(01,01,01,07,07,07),
  "Variable" = c("language","edition","author","language","edition","author"),
  "Values" = c("Java","third","Herbert Schmidt","C++","second","E.Balagurusamy")
)
tall_books
```

CHUNK 27: Convert tall data to wide
```{r}
# converts to wide
wide_books <- pivot_wider(tall_books,names_from=Variable, values_from = Values)
wide_books
```

CHUNK 28: Convert wide data to tall
```{r}

tall_again <- pivot_longer(wide_books,c("language","edition","author"),names_to="Variable",values_to="Values")
tall_again
```

CHUNK 29: Exercise 2.2.5. Convert the tall_measure.csv data set to a wide data set. Why is there a missing data point after converted when there wasn't one before?
```{r}
long_data <- read_csv("tall_measure.csv")

```

CHUNK 30: Exercise 2.2.5: Solution
```{r}
long_data <- read_csv("tall_measure.csv")

pivot_wider(long_data,names_from=Month,values_from=Measurement)

# There is one fewer observation in the tall data set for
# Steve than there is for Tanya, so the extra observation
# is missing for Steve
```

CHUNK 31: Create a data frame
```{r}

library(tidyverse)
my_dat <- tibble(
  "first_variable" = c(1, 2, 3),
  "second_variable" = c(6, 7, 8)
)
my_dat
```

CHUNK 32: Write data to text file
```{r}

write_delim(my_dat, "datafile_space.txt") # Space delimiters is the default, see help on using the function to use other characters as delimiters

# Comma delimited
write_csv(my_dat, "datafile_comma.csv")
```

CHUNK 33: Appending a file
```{r}

new_dat <- tibble(
  "first_variable" = c(4, 5),
  "second_variable" = c(9, 10)
)
write_csv(new_dat, "datafile_comma.csv", append = TRUE)

read_csv("datafile_comma.csv")
```